\documentclass[11pt]{article}
\usepackage{fullpage}

\begin{document}

\title{Homework 2\\
       CSc 72700: Analysis of Algoritms\\
       CUNY Graduate Center, Fall 2001 }
\date{Due Wednesday, 12 September}
\author{}
\maketitle

See the guidelines on the webpage for details about submitting homework.
(If turning your homework in electronically, you can mail it directly
to the grader at: {\tt ivm3@columbia.edu}.)

\section*{Practice Problems}

The problems in this section {\bf are not to be submitted}.  They are to
help you understand the material, and some will appear on exams.

\begin{itemize}
    \item Problem 4-7 on p 75 (in the second edition: 4-6 on p 87).
    \item Exercise 7.1-1 on p 142 (in the second edition: 6.1-1 on p 129).
    \item Exercise 7.1-2 on p 142 (in the second edition: 6.1-2 on p 129).
    \item Exercise 7.2-1 on p 144 (in the second edition: 6.2-1 on p 132).
    \item Exercise 7.2-4 on p 144 (in the second edition: 6.2-4 on p 132).
    \item Exercise 7.3-2 on p 147 (in the second edition: 6.3-3 on p 135).
    
\end{itemize}

\section*{Graded Problems}

These problems will be graded and should be submitted, following the
guidelines on the webpage.

\begin{enumerate}
    \item Exercise on worst case running time for Heapify,
	7-2.5 on p 144 (in the second edition: 6.2-6 on p 132).
    \item Exercise 7.3-3 on p 147 (in the second edition: 6.3-3 on p 135).
    \item Problem on building an heap using insertion,
	7-1 on p 152 (in the second edition: 6-1 on p 142).

\end{enumerate}

\end{document}
