\documentclass[11pt]{article}
\usepackage{fullpage}

\begin{document}

\title{Homework 3\\
       CSc 72700: Analysis of Algoritms\\
       CUNY Graduate Center, Fall 2001 }
\date{Due Wednesday, 3 October}
\author{}
\maketitle

See the guidelines on the webpage for details about submitting homework.
(If turning your homework in electronically, you can mail it directly
to the grader at: {\tt ivm3@columbia.edu}.)

\section*{Practice Problems}

The problems in this section {\bf are not to be submitted}.  They are to
help you understand the material, and some will appear on exams.  

\begin{itemize}
    \item Exercise 8.1-1 on p 155 (in the second edition: 7.1-1 on p 148).
    \item Exercise 8.1-2 on p 155 (in the second edition: 7.1-2 on p 148).
    \item Exercise 8.3-1 on p 162 (in the second edition: 7.3-1 on p 154).
    \item Exercise 12.2-2 on p 226 (in the second edition: 11.2-2 on p 229).
    \item Problem 12-1 (Longest probe for bounding hasing) on p 241 
	(in the second edition: 11-1 on p 250).
    
\end{itemize}

\section*{Graded Problems}

These problems will be graded and should be submitted, following the
guidelines on the webpage.

\begin{enumerate}
    \item Exercise on running time for quicksort when all elements have the
	same value,
	8.2-1 on p 160 (in the second edition: 7.2-2 on p 153).
    \item Exercise on worst case searching when $|u|>nm$,
	12.2-6 on p 226 (in the second edition: 11.2-5 on p 229).
    \item Implement {\bf  both} the regular and randomized version of
	quicksort.  Run the sorts on a list of 1000 randomly generated
	numbers, and on the list: {\tt 1000, 999, 998,..., 3,2,1}.
	What difference in performance did you notice?  Give an possible
	explanation for any difference you noticed.  
	[To measure 
	the time performance, you could time it using profiling tools
	on your system, or another way would be to add in a counter or
	print statements and count the number of calls.]

	You should submit your programs, the timing results for both
	data sets on both programs, and a discussion of their relative
	performance.

\end{enumerate}

\end{document}
