\documentclass[11pt]{article}
\usepackage{fullpage}

\begin{document}

\title{Homework 4\\
       CSc 72700: Analysis of Algoritms\\
       CUNY Graduate Center, Fall 2001 }
\date{Due Friday, 19 October}
\author{}
\maketitle

See the guidelines on the webpage for details about submitting homework.
(If turning your homework in electronically, you can mail it directly
to the grader at: {\tt ivm3@columbia.edu}.)

\section*{Practice Problems}

The problems in this section {\bf are not to be submitted}.  They are to
help you understand the material, and some will appear on exams.

\begin{itemize}
    \item Exercise 13.1-1 on p 246 (in the second edition: 12.1-1 on p 256).
    \item Exercise 13.2-1 on p 249 (in the second edition: 12.2-1 on p 259).
    \item Exercise 13.3-1 on p 253 (in the second edition: 12.3-1 on p 264).
    \item Exercise 14.1-4 on p 265 (in the second edition: 13.1-6 on p 277).
    \item Exercise 14.2-2 on p 267 (in the second edition: 13.2-2 on p 278).
    \item An {\bf AVL} tree is a binary search tree in which the subtrees of
	a node differ in height by at most one.  Insertions and deletions 
	are performed as in binary search trees and then followed by  
	rotations to preserve balance (see links on webpage for more details
	or 13-3, p 296 in the second edition).
	
	Construct a binary search tree by introducing the following keys 
	in the given order: 2, 6, 10, 8, 3, 4, 5.  Using the same sequence 
	of keys, construct an AVL tree that uses rotations to keep balance
	after every insertion.
\end{itemize}

\section*{Graded Problems}

These problems will be graded and should be submitted, following the
guidelines on the webpage.

\begin{enumerate}
    \item Average node depth in a randomly built binary search tree,
	13-3 on p 260 (in the second edition, 12-3 on p 270).
    \item Number of different binary trees,
	13-4 on p 262 (in the second edition, 12-4 on p 271).

	Note, there is a typo in the first edition.  In part b), the
	formula should read:
	$$
		f(x) = \sum_{k=0}^{\infty} \frac{f^{(k)}(a)}{k!} {(x-a)}^k
	$$
    \item AVL Trees (see webpage or 13-3, p 296 second edition for 
	defintions)

	\begin{enumerate}
	    \item Trees of height $O(\lg n)$ are called {\bf balanced}.
		Show AVL trees are balanced.
	    \item What is the most number of rotations that must be 
		performed when inserting to maintain balance?  
		What is the most number 
		of rotations that must be performed, when deleting?
	    \item What is the running time of the operations: insert,
		delete, and search for AVL trees?
	\end{enumerate}


\end{enumerate}

\end{document}
