\documentclass[11pt]{article}
\usepackage{fullpage}

\begin{document}

\title{Homework 8\\
       CSc 72700: Analysis of Algoritms\\
       CUNY Graduate Center, Fall 2001 }
\date{Due Wednesday, 12 December 2001}
\author{}
\maketitle

See the guidelines on the webpage for details about submitting homework.
(If turning your homework in electronically, you can mail it directly
to the grader at: {\tt ivm3@columbia.edu}.)

\section*{Practice Problems}

The problems in this section {\bf are not to be submitted}.  They are to
help you understand the material, and some will appear on exams.

\begin{itemize}
    \item Bottleneck TSP, 37.2-3 on p 974
	(in the second edition, 35.2-3, p 1033).
    \item Non-crossing tours, 37.2-4 on p 974
	(in the second edition, 35.2-4, p 1033).
    \item In the Albers notes, three algorithms for the on-line 
	$k$-paging problem are suggested: LFU (Least Frequently Used),
	LRU (Least Recently Used), and FIFO (First In First Out).  The
	first one is considered in a graded problem below.  In the notes,
	the latter two are shown to $k$-competitive.  Are they $m$
	competitive for $m < k$?
    \item Hiring exactly once,
	in the second edition (handout), 5.2-1, p 98.
    \item Inversions,
	in the second edition (handout), 5.2-5, p 99.
    \item Permute with all,
	in the second edition (handout), 5.3-3, p 105.
\end{itemize}

\section*{Graded Problems}

These problems will be graded and should be submitted, following the
guidelines on the webpage.

\begin{enumerate}
    \item Bin Packing\\
	Problem 37-1 on p 983
	(in the second edition, 35-1 on p 1049).
    \item Deterministic $k$-paging algorithms\\
	Consider a machine with $k$ fast memory locations and a 
	large slow memory.  The input is a sequence of page requests.
	If the page is fast memory, the request is satisfied.  If it
	is not, a {\em page fault} occurs while some page is removed
	from main memory and the desired page is loaded into the 
	empty spot.  We measure the cost of an algorithm on an input
	sequence $\sigma$ by the number of page faults required to 
	service all the page requests
	(see Albers notes, p 1-4 for more details).

	Let {\bf LFU} (Least Frequently Used) be the algorithm that on
	a page fault, evicts the page in memory that was used the 
	least frequently.
	\begin{enumerate}
	    \item Show that for each $k$, there exists an instance
		$\sigma$ such that
		$$
		    C_{LFU}(\sigma) > 2 \cdot C_{MIN}(\sigma)
		$$
	    \item Fix the $k$, the number of fast memory registers.
		Show that for each $m > 0$, there exists an instance
		$\sigma$ such that
		$$
		    C_{LFU}(\sigma) > m \cdot C_{MIN}(\sigma)
		$$
	    \item Conclude that the algorithm LFU is not competitive
		(i.e. not $m$-competitive for any $m$).
	\end{enumerate}
	
    \item The hat-check problem\\
	In the second edition (handout), 5.2-4, p 98.
	
\end{enumerate}

\end{document}
