\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{amssymb}

\textwidth = 6.5 in
\textheight = 9 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.0 in
\parskip = 0.2in
\parindent = 0.0in

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}

\begin{document}

\begin{center}
	Exam 2\\
        	CSc 75010: Theoretical Computer Science\\
        	Graduate Center of CUNY\\
	22 November 2002\\
        	(All Questions)
\end{center}

Do five of the following six problems.  Write each answer on a separate
piece of paper.


\begin{enumerate}
%1.  a)  Give a TM definition that doubles a number, subtracts one in unary, etc.
%	b) Define computable function and give an example of a non-computable function.
\item
\begin{enumerate}
	\item Define {\em decidable set} and give an example of a set that is not decidable.
	\item Assume that the alphabet and tape alphabet are: 
		$\Sigma = \Gamma = \{0,1\}$.
		Give an implementation level description of a Turing machine that
		decides the language: 
		$$
			\{w\mid \mbox{$w$ contains an equal number of 0's and 1's}\}
		$$

	\item Define {\em computable function} and give an example of a function that is not computable.
	\item Assume that the alphabet and tape alphabet are:  $\Sigma = \Gamma = \{0,1\}$.
		Give an implementation level description of a Turing machine that
		copies the input string on the tape.  That is, if the input to the machine
		is the string $w$, the output is $ww$.

	\item Define {\em decidable set} and give an example of a set
		that is not decidable.
	\item Assume that the alphabet and tape alphabet are: $\Sigma = \Gamma = \{0,1\}$.
		Give an implementation level description of a Turing machine that
		decides the language: 
		$$
			\{w\mid \mbox{$w$ is a palindrome}\}
		$$

	\item Define {\em decidable function} and give an example of a function
		that is not decidable.
	\item Assume that the alphabet and tape alphabet are: 
		$\Sigma = \Gamma = \{0,1\}$.
		Give an implementation level description of a Turing machine that
		takes as input a number in binary representation and mulitplies it
		by 2 (that is, shifts it to the right).

\end{enumerate}

%2.  Show something countable, and related item uncountable: $\sigma^*$ and all strings;
%	rationals and reals;
\item 
\begin{enumerate}
	\item Show that for any finite set $\Sigma$, the set of all finite strings of $\Sigma$, 
		$\Sigma^*$ is countable.
	\item Let $\Sigma = \{0,1\}$.  Show that the set of all infinite strings over $\Sigma$ is 		uncountable.
	
	\item Show that the set of all positve rational numbers is countable.
	\item Show that the set of all positve real numbers is uncountable.
	
\end{enumerate}
%3.  a) Use the diagonalization method to show K is undecidable.
%	b) Use this to show Tot, Min, Empty, Const is decidable.
\item 
\begin{enumerate}
	\item Show that the Halting Problem is undecidable, using the diagonalization method.
	\item Show that the set of all finitely defined computable functions (that is, the set of 
		all functions that diverge for all but a finis not decidable.

	\item Show that the Halting Problem is undecidable, using the diagonalization method.
	\item Show that the set of all constant computable functions is not decidable.
	
	\item Show that the Halting Problem is undecidable, using the diagonalization method.
	\item Show that the set of all totally undefined computable functions (that is the set of
		all functions that diverge for all inputs) is not decidable.
		
	\item Show that the Halting Problem is undecidable, using the diagonalization method.
	\item Show that the set of all totally defined computable functions is not decidable.
	
\end{enumerate}
%4.  a) Show that the set of Turing-recognizable/decidable languages is closed under union,
%	concatenation, star, intersection.
%	b) Show if A and $\bar{A} are Turing-recognizable, then A is decidable.
\item
\begin{enumerate}
	\item Show that the set of decidable languages is closed under concatenation.
	\item Show that if $A$ and $\bar{A}$ are Turing-recognizable, then $A$ is decidable.

	\item Show that the set of Turing-recognizable languages is closed under star.
	\item Show that if $A$ is decidable, then $A$ and $\bar{A}$ are Turing-recognizable.	
	\item Show that the set of decidable languages is closed under intersection.
	\item Show that if $A$ and $\bar{A}$ are Turing-recognizable, then A is decidable.
	
	\item Show that the set of decidable languages is closed under union.
	\item Show that if $A$ and $\bar{A}$ are Turing-recognizable, then A is decidable.
	
	\item Show that the set of Turing-recognizable languages is closed under concatenation.
	\item Show that if $A$ is decidable, then $A$ and $\bar{A}$ are Turing-recognizable.
	
\end{enumerate} 
%5.  a) State PCP, 
%	b) show decidable over $\Sigma = \{1\}$.
%	or: State and prove Rice's Theorem.
\item 
\begin{enumerate}
	\item State the Post Correspondence Problem (PCP).
	\item Show that PCP is decidable over the alphabet $\Sigma = \{1\}$.
	
	\item State Rice's Theorem.
	\item Prove Rice's Theorem.
\end{enumerate}

%6.  $\leq_m$ questions: 5.3: if $A\leq_m B$ and $B$ regular does that imple A is regular?
%	5.6: $\leq_m$ is transitive, 5.7:  A r.e. and $A \leq_m \bar{A}, then A is decidable,
%	fact:  $A \leq_m B$ implies $\bar{A} \leq_m \bar{B}$.
%	b) Turing Reducibility: 6.1: $A\leq_T B$ and $B\leq_T C$ implies $A\leq_T C$,
%	6.7: Show $\forall A, B \exists J$ such that $A\leq_T J$ and $B\leq_T J$,
\item 
\begin{enumerate}
	\item If $A\leq_m B$ and $B$ regular does that imple A is regular?  Why or why not?
		Justify your answer.
	\item Show for all $A$, $B$, there exists a set  $J$ such that 
		$A\leq_T J$ and $B\leq_T J$.
		
	\item Show that $\leq_m$ is transitive.
	\item Show for all $A$, $B$, there exists a set  $J$ such that 
		$A\leq_T J$ and $B\leq_T J$.	
		
	\item Show that if $A$ Turing-recognizable and $A \leq_m \bar{A}$, then A is decidable.
	\item Show that if $A\leq_T B$ and $B\leq_T C$ implies $A\leq_T C$.
	
	\item Show that for all $A$ and $B$,  $A \leq_m B$ implies $\bar{A} \leq_m \bar{B}$.
	\item Show for all $A$, $B$, there exists a set  $J$ such that 
		$A\leq_T J$ and $B\leq_T J$.	

\end{enumerate}

\end{enumerate}

\end{document}
