\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{amssymb}

\textwidth = 6.5 in
\textheight = 9 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.0 in
\parskip = 0.2in
\parindent = 0.0in

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}

\begin{document}

\begin{center}
	Exam 1\\
        	CSc 75010: Theoretical Computer Science\\
        	Graduate Center of CUNY\\
	4 October 2002\\
        	(Sample Exam)
\end{center}

Do five of the following six problems.  Write each answer on a separate
piece of paper.


\begin{enumerate}
%Define the following terms:
	\item Define the following terms:
     	   \begin{enumerate}
		\item regular expression
		\item Pigeonhole Principle
		\item stack
		\item Given a string $s$, define $|s|$
		\item Given finite sets $\Sigma_1, \Sigma_2$, 
			define $\Sigma_1 \circ \Sigma_2$
	   \end{enumerate}
%What is wrong with this proof?
   	\item Find the error in the following proof that all horses are the same color.\\
		CLAIM: In any set of $h$ horses, all horses are the same color.\\
		PROOF: By induction on $h$.\\
		{\bf Basis:} For $h=1$.  IN any set containing just one horse, all horses clearly
			are the same color.
		{\bf Induction step:} For $k\geq 1$, assume that the claim is true for $h=k$ and
			prove that it is true for $h=k+1$.   Take any set $H$ of $k+1$ horses.  We
			show that all the horses in this set are the same color.  Remove one horse
			from this set to obtain the set $H_1$ with just $k$ horses.  By the induction 
			hypothesis, all the horses in $H_1$ are the same color.  Now replace the 
			removed horse and remove a different one to obtain the set $H_2$.  By
			the same argument, all horses in $H_2$ are the same color.  Therefore,
			all the horses in $H$ nyst be the same color, and the proof is complete.
%Build DFA's
   	\item Give the state diagrams of NFAs recognizing the following languages.  In all
	    cases the alphabet is $\Sigma = \{a,b,c,d,\ldots,x,y,z\}$, the 26 lowercase letters.
	    \begin{enumerate}
		\item $\{w \mid w \mbox{ contains the substring $help$}\}$
		\item $\{w \mid w \mbox{ is of length at least 2 and an even numbers of $z$'s}\}$
	    \end{enumerate}
%Show closure under one of the regular operators
   	\item  Prove that the class of regular languages is closed under the union operator.
%Show the language is not regular
   	\item Prove that the following language is not regular:
		$$
			\{ w \mid w \in \{0,1\}^* \mbox{ is not a palindrome}\}
		$$
%Given context-free grammar, what is the lang, make PDA
   	\item Give context-free grammars generating the following languages:
	  \begin{enumerate}
		\item $\{w \mid \mbox{the length of $w$ is odd and its middle symbol is $0$}\}$
		\item The complement of the language $\{a^n b^n \mid n\geq 0\}$
	  \end{enumerate}
\end{enumerate}



 \end{document}