\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\pagestyle{empty}

\textwidth = 6.5 in
\textheight = 9 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.0 in
\parskip = 0.2in
\parindent = 0.0in

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}

\begin{document}

{\bf
\begin{tabular}{c}
	Name: \underline{\hspace{2.5in}}\\ \\
	{\large Final Exam}\\
        	CSc 75010: Theoretical Computer Science\\
        	Graduate Center of CUNY\\
	13 December 2002\\
        	%(All questions)
\end{tabular}
}
\hfill
\begin{tabular}{|l|c|}
\hline
Question 1 & \mbox{\hspace{.5in}}\\
\hline        	
Question 2 & \\
\hline
Question 3 & \\
\hline
        	Question 4 & \\
\hline
        	Question 5 & \\
\hline
        	Question 6 & \\
\hline\hline
         Total & \\
\hline
\end{tabular}

Do five of the following six problems.  Write each answer on a separate
piece of paper.


\begin{enumerate}

\item Let $\Sigma = \{0,1\}$ be a finite alphabet.
	\begin{enumerate}
	\itemsep = 0pt
		\item Define {\em regular language}.
		\item Show the following language is regular over $\Sigma$:
		$
			\{ w \mid \mbox{ $w$ ends in a $0$} \}
		$
		\item Show the following language is not regular over $\Sigma$:
		$
			\{ w \mid w = 0^n1^n, n \geq 0 \}
		$
	\end{enumerate}

\item For each of the following statements, state whether it is true or false.
	Justify your answer by providing a proof sketch or counterexample.
	\begin{enumerate}
	\itemsep= 0pt
		\item Every regular language is context-free.
		\item Every regular language is Turing-recognizable.
		\item Every decidable language is regular.
	\end{enumerate}

\item 
	\begin{enumerate}
	\itemsep= 0pt
		\item State the Halting Problem.
		\item Prove the Halting Problem is Turing-Recognizable.
		\item Prove the Halting Problem is not decidable.
	\end{enumerate}

\item 
	\begin{enumerate}
		\itemsep=0pt
		\item State the Post Correspondence Problem (PCP).
		\item Is the PCP undecidable for all alphabets?
			Why or why not?  If yes, sketch a proof of the undecidability.  If no, 
			give an example of an alphabet over which PCP is decidable.  Justify
			your answer.
	\end{enumerate}

\item 
	\begin{enumerate}
		\itemsep=0pt
		\item State the Recursion Theorem.
		\item Given a Turing Machine $M$, the {\em length} of the description of
			$<\!\!\!M\!\!\!>$ is the number of symbols in the string describing $M$.  $M$
			is {\em minimal} if there is no Turing Machine equivalent to $M$ with
			a shorter description.  Show
			$$
				\mbox{MIN}_{TM} = \{<\!\!\!M\!\!\!> \mid \mbox{$M$ is a minimal TM} \}
			$$ 
			is not Turing-recognizable.
	\end{enumerate}

\item For the following sets, state whether the set is decidable or Turing-recognizable
	(or both).  Justify your answer.
	\begin{enumerate}
		\itemsep=0pt
		\item $E_{TM} = \{<\!\!\!M\!\!\!> \mid \mbox{$M$ is a TM and $L(M) = \emptyset$}\}$
		\item ${Th}({\bf N}, +)$
		\item ${Th}({\bf N}, +, \times)$			
	\end{enumerate}


\end{enumerate}

\end{document}
