\documentstyle[11pt]{article}
\pagestyle{empty}
\parskip 1em
\topmargin 1pt \headheight 1pt \headsep 2pt \footskip 1pt 
\oddsidemargin -10pt \evensidemargin -10pt \parindent 0pt
\begin{document}
\input{epsf}

Computer Science 125 - 002 \hspace*{1in} November 3, 1998 \hspace*{1in} Second Exam\\

1. In the following, you are to define a portion of a {\tt Circle}
class (that does not invlolve any graphics).
Include the following items:
\begin{itemize}
\item data members for the x and y coordinates
of the center and the radius.
\item default constructor which creates a circle with
center at the origin and unit radius.
\item a constructor which takes a supplied x and y coordiantes
for the center and a radius.
\item a method which computes and returns the area.
\end{itemize}
[24 points]





2. The following technique was used by a software
company:

Users who installed a product from the company
were invited to run a supplied program to
generate a mailing label 
for product registration.  The
software scanned the user's disk for all products which
were produced by the company 
and included the encoded inventory information
on the label (users were not informed of its exact behavior).  
The company then used the
information to trace and pursue people who 
had illegally installed its software.

Do you think that the above practice was a violation of
privacy rights?  Support your position.  [12 points]

3. The following function does a linear search of
in integer array through a specified number of
elements for a value which is passed as a parameter and
returns the index where found or -1 if not found.


\begin{verbatim}
    public int LSearch(int[] a, int count, int value)
    {
        int index = 0;
        while (index < count)
        {
            if (a[i] == value)
                return index;
            else
                index++;
        }
        return -1;
    }
\end{verbatim}

If the arrary {\tt a} is in ascending order, then
the above can be improved slightly so that 
unsuccessful searches are generally somewhat better.
Make the required changes (DO NOT turn it into
a binary search). [10 points]



4. Given the following problem statement, circle the words
(or phrases) which are candidates for the major classes
in the problem.  Also underline items which
are good candidates for methods and for each such item,
draw an arrow to the class in which it should be placed.
Problem statement: [14 points]

\begin{quote}

Design a program which performs additions and multiplications
with polynomials.  A polynomial will consist of one or
more terms where a term consists of a nonzero coefficient
and a nonnegative integer exponent.

\end{quote}

5. Testing is an important activity in system development [8 points each].

(a) Differentiate between black-box and white-box unit testing.

(b)  Each of the methods in (a) tend to expose different
kinds of errors.  For one of the techniques, indicate a 
(type of) problem which it 
may be more effective at exposing than the other.  


6. Write a function named {\tt Transpose} which takes a two-dimensional
double array
parameter, creates,  and returns the transpose of its
parameter.  The {\em transpose} of an MxN array is
an NxM array for which an entry in the (i,j) position comes
from the (j,i) position from the other matrix.  For example,
the transpose of
\[ \left (
\begin{array}{cc}
1 & 2\\
3 & 4\\
5 & 6\\
\end{array}
\right )
\ \ \ is\ \ \
 \left (
 \begin{array}{ccc}
 1 & 3 & 5\\
 2 & 4 & 6\\
 \end{array}
 \right ) \]

 [14 points]


 7.  Draw a picture showing all of the storage cells associated with
 the statement: [10 points]

 {\tt int[][][] a = new int[2][3][2];}
 
\end{document}

