\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\pagestyle{empty}

\textwidth = 6.25 in
\textheight = 9 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.0 in
\parskip = 0.2in
\parindent = 0.0in

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}

\begin{document}

{\bf\large
\begin{center}
	{\large First Exam}\\
        	Theoretical Computer Science\\
        	Graduate Center\\
	City University of New York\\
	10 June 2003
\end{center}
}

%\begin{tabular}{|l|c|}
%\hline
%Question 1 & \mbox{\hspace{.5in}}\\
%\hline        	
%Question 2 & \\
%\hline
%Question 3 & \\
%\hline
%        	Question 4 & \\
%\hline
%        	Question 5 & \\
%\hline
%        	Question 6 & \\
%\hline\hline
%         Total & \\
%\hline
%\end{tabular}

Do three of the following four problems.  Write each answer  in a separate
blue book.


\begin{enumerate}

%Modified from final
\item Let $\Sigma = \{0,1\}$ be a finite alphabet.
	\begin{enumerate}
	\itemsep = 0pt
		\item Define {\em regular language}.
		\item Show the following language is regular over $\Sigma$:
		$
			L = \{ w \mid \mbox{$w$ contains the substring $1001$} \}
		$
		\item Define a Turing machine that accepts the language $L$ above.
		\item Show the following language is not regular over $\Sigma$:
		$
			\{ w \mid w = 0^m1^n, m \geq n \}
		$
	\end{enumerate}


\item 
	\begin{enumerate}
	\itemsep= 0pt
		\item State the Halting Problem.
		\item Prove the Halting Problem is Turing-Recognizable (i.e. recursively enumerable).
		\item Prove the Halting Problem is not decidable.
	\end{enumerate}

\item 
	\begin{enumerate}
		\itemsep=0pt
		\item State the Recursion Theorem.
		\item Prove the Recursion Theorem.
		\item Prove the following fixed point theorem:\\
			Let $t:\Sigma^* \rightarrow \Sigma^*$ be a computable function.
			Then there is a Turing machine $F$ wherein $t(<F>)$ describes a
			Turing machine equivalent to F.
			%Thm 6.6, p 203
	\end{enumerate}

%Too hard:
%\item 
%	\begin{enumerate}
%		\itemsep=0pt
%		\item State the Recursion Theorem.
%		\item Given a Turing Machine $M$, the {\em length} of the description of
%			$<\!\!\!M\!\!\!>$ is the number of symbols in the string describing $M$.  $M$
%			is {\em minimal} if there is no Turing Machine equivalent to $M$ with
%			a shorter description.  Show
%			$$
%				\mbox{MIN}_{TM} = \{<\!\!\!M\!\!\!> \mid \mbox{$M$ is a minimal TM} \}
%			$$ 
%			is not Turing-recognizable.
%	\end{enumerate}

\item For the following sets, state whether the set is decidable or Turing-recognizable
	(or both).  Justify your answer.
	\begin{enumerate}
		\itemsep=0pt
		\item $E_{TM} = \{<\!\!\!M\!\!\!> \mid \mbox{$M$ is a TM and $L(M) = \emptyset$}\}$
		\item A set $A$, where $A \leq_m B$ and B is decidable.
		\item A set $A$, where $\bar{A} \leq_m A$ and $\bar{A}$ is Turing-recognizable.
		\item ${Th}({\bf N}, +)$
		\item ${Th}({\bf N}, +, \times)$			
	\end{enumerate}


\end{enumerate}

\end{document}
